---
title: "Prediction of 2024 US election ..."
author: 
  - Colin Sihan Yang
  - Lexun Yu
  - Siddharth Gowda
thanks: "Code and data are available at: [https://github.com/yulexun/uselection](https://github.com/yulexun/uselection)."
date: today
date-format: long
abstract: "We forecast the winner of the 2024 US presidential election using “poll-of-polls” by building a linear model."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(here)
library(tibble)
library(knitr)
library(maps)
library(mapproj)
light_blue <- "#ADD8E6"
dark_blue <- "#6688CC"
light_red <- "#FF7777"
dark_red <- "#B22222"
purple <- "#9370DB"
```


# Introduction {#sec-intro}
Election result forecasting has become an essential tool for analysts in political science and the public to predict the outcome of democratic process, such as the presidential election in the United States. Traditionally, individual polls have been used as a snapshot of voter sentiment, but they only reflect temporary changes in the performance of contestants, instead of a precise estimation of the election result. As discussed by @pasek and @blumenthal, the aggregation of multiple polls, or "poll-of-polls," has become a popular technique to reduce individual survey errors and provide more accurate election forecasts. However, the traditional poll aggregation does not reflect dynamics of an election, especially with real-time changes and the introduction of new data. This creates a gap for a more adaptable model to predict the election result based on both polling data and additional variables, such as historical data and economic indicators.

This paper fills the gap by building a hybrid election forecasting model following the strategies mentioned by @pasek. As @pasek described in their article, aggregation involves determining which surveys are worth including, as well as selecting, combining and averaging results from multiple polls to reduce individual biases and errors. Prediction modeling adds other data to the model that predicts election outcomes based on current dynamics. Hybrid models like the Bayesian approach incorporates prior beliefs based on historical data or expert knowledge and new evidence like economic updates to dynamically adjust the forecast as the campaign progresses. 

In this paper, we aim to predict the 2024 us election result with the hybrid election forcasting model. We incorporate aggregation by filtering the polls on @fivethirtyeight by numeric grade that indicates pollster’s reliability, prediction that incorporates social and economic indicators including unemployment rates and abortion rates, and hybrid approaches that leverages Bayesian techniques which combines historical data such as the 2016 election data, allowing for a dynamic prediction of the U.S. presidential election. 

The estimand for this research paper is the predicted support percentages for Kamala Harris and Donald Trump. The prediction is based on quantifying various polling factors, including sample size, poll scores, and transparency scores, which are used as predictors.

The results of this model indicate a more stable and accurate forecast compared to traditional aggregation methods alone, [update this …]

The remainder of this paper is structured as follows: [update this …]


# Data {#sec-data}

## Overview

For the data we used in this analysis about the polling result for Kamala Harris and Donalad Trump in 2024 USA president election. \
- **response variable:** pct(pct: The percentage of the vote or support that the candidate received in the poll) \
- **numeric predictor:**\
 sample size(sample_size: The total number of respondents participating in the poll) \
 timegap(the time gap between the poll start date and the real election date i.e timegap = real US election date - poll start date) \
 pollscore(A numeric value representing the score or reliability of the pollster in question) \
- **categorical predictor**
state(The U.S. state where the poll was conducted or focused) \
methodology(The method used to conduct the poll)\



## Measurement

In this dataset, each row represents a polling question that records the variables of interest. Each entry allows us to explore the real-world relationships between polling factors and the support percentage (`pct`) for the candidates Kamala Harris and Donald Trump. This dataset enables an analysis of how various polling characteristics influence the reported support levels for the candidates we are focused.


## Clean Data
The data cleaning process involves several steps to ensure the quality and relevance of the polling data. First, we filter the dataset to retain only poll results with a numeric grade of 2.7 or higher, indicating that the polls are considered reliable. Next, we address missing values in the state attribute: polls with NA in the state column are considered national polls.

We then create a new attribute, days_taken_from_election, which represents the time gap between the poll's start date and the actual U.S. election date. Additionally, we filter the dataset to include only polls conducted after July 21, 2024, the date when Kamala Harris declared her candidacy. Finally, we remove any remaining rows that contain missing values to ensure a clean dataset.

```{r}
#| label: tbl-cleaned-data
#| tbl-cap: Sample of cleaned US election data 
#| echo: false
#| warning: false
cleaned_data = read_csv(here("data/02-analysis_data/cleaned_data.csv"))

cleaned_data |>
  select(pct, sample_size, pollscore, days_taken_from_election, state, methodology, candidate_name) |>
  head(6) |>
  kable(
    col.names = c("pct", "sample_size", "pollscore", "days_taken_from_election","state", "methodology", "Candidate Name"),
    booktabs = TRUE
    ) 
```

### States included in anaylsis

After the data cleaning process, 21 states had no polling data. A table showing the number of polls for each state, including those without any polls, is provided in @tbl-state-polls-count.

This absence of polling data is not a significant concern due to the structure of the United States Electoral College (explained in detail in the Appendix). The states lacking polling data have consistently followed historical voting patterns, so predicting the winning candidate in those states is unnecessary

## Basic Statistics Summary for Data

```{r}
#| label: fig-avg-pct-state
#| fig-cap: Historical State Voting Trends Are Maintined in 2024
#| echo: false
#| warning: false
#| 
cleaned_data <- cleaned_data %>% mutate(candidate_name =
  if_else(!(candidate_name %in% c("Kamala Harris", "Donald Trump")), "Other", candidate_name))

cleaned_data %>%
  filter(candidate_name %in% c("Kamala Harris", "Donald Trump")) %>%
  group_by(state, candidate_name) %>%
  summarise(avg_pct = mean(pct, na.rm = TRUE), .groups = "drop") %>%
  ggplot(mapping = aes(x = avg_pct, y = state, fill = candidate_name)) + 
  geom_col(position = "stack") +
  theme_minimal() +
  labs(
    x = "Average Polled Support (%)", 
    y = "State",
    fill = "Candidate"
  ) +
  scale_fill_manual(values = c("Kamala Harris" = light_blue, "Donald Trump" = light_red)) +
  theme_minimal() +
  theme(legend.position = "bottom") + 
  theme(axis.text.y = element_text(size = 8))

```
In figure @fig-avg-pct-state, historically Democratic are polling for Kamala and historically Republican states are polling for Trump. Similarly, historically swing states also appear to be close, for instance Michigan (46.9% Trump, 47.5% Harris), Nevada (47.7% Trump, 47.9% Harris), and Pennsylvania (46.9% Trump, 48.2% Harris) all are close to an even split.

```{r}
#| label: fig-avg-pct-methodology
#| fig-cap: Most Metholodgies Do Not Bias Towards a Candidate
#| echo: false
#| warning: false
#| 
cleaned_data <- cleaned_data %>% mutate(candidate_name =
  if_else(!(candidate_name %in% c("Kamala Harris", "Donald Trump")), "Other", candidate_name))

cleaned_data %>%
  filter(candidate_name %in% c("Kamala Harris", "Donald Trump")) %>%
  filter(!is.na(methodology)) %>% 
  ggplot(mapping = aes(x = methodology, y = pct)) + 
  geom_boxplot(width = 0.6, position = position_dodge(width = 0.8), outlier.shape = NA) +
  coord_flip() +
  theme_minimal() +
  labs(
    x = "Methodology",
    y = "Polled Support (%)",
    fill = "Candidate"
  ) +
  scale_fill_manual(values = c("Kamala Harris" = light_blue, "Donald Trump" = light_red)) +
  theme(legend.position = "bottom") + 
  theme(axis.text.y = element_text(size = 8))

```
Figure @fig-avg-pct-methodology shows that polls utilizing multiple communication methods to reach voters tend to have lower interquartile ranges (IQRs) in their boxplots compared to those that rely on only one or two communication methods.

## Relationship Between Variables

```{r}
#| label: fig-pollscore-samplesize
#| fig-cap: More Reliable Pollsters Have Larger Sample Sizes in their Polls
#| echo: false
#| warning: false

# removing outliers and na's
graph_data <- cleaned_data %>% 
  filter(!is.na(pollscore) & !is.na(sample_size)) 

outlier_top <- (quantile(graph_data$sample_size, 0.75) + 
                  IQR(graph_data$sample_size)*1.5)

outlier_bottom <- (quantile(graph_data$sample_size, 0.25) - 
                  IQR(graph_data$sample_size)*1.5)

graph_data <- cleaned_data %>%
  filter(sample_size < outlier_top & sample_size > outlier_bottom)

graph_data %>% filter(!is.na(pollscore) & !is.na(sample_size)) %>% 
  ggplot(aes(x = pollscore, y = sample_size)) +
  geom_point(alpha = 0.33) + geom_smooth(method = "lm", se = FALSE, color = light_red) +
  labs(x = "Poll Score", y = "Poll Sample Size")

# this model is really here just for explain, this realtionship should not
# be treated as lienar
# cor(graph_data$sample_size, graph_data$pollscore)
# lm_score_sample <- lm(sample_size ~ pollscore, data = graph_data)
# summary(lm_score_sample)
```
Figure @fig-pollscore-samplesize illustrates a weak positive correlation between a pollster's pollscore and the sample size of their poll. The figure also shows that most polls have a sample size around 800 to 1200 participants. It is also important to note that a few polls with exceptionally large sample sizes were excluded from the graph due to their status as clear outliers.

```{r}
#| label: fig-pollsamplesize-canidate-pct
#| fig-cap: The Sample Size of a Poll does not impact a Candidate's Voting Percentage
#| echo: false
#| warning: false

cleaned_data %>% 
  filter(!is.na(sample_size) & !is.na(pct)) %>% 
  ggplot(aes(x = sample_size, y = pct)) +
  geom_point(alpha = 0.4) + 
  geom_smooth(method = "lm", se = FALSE, color = light_red) +
  labs(x = "Sample Size of the Poll",
       y = "Candidate Percentage") + facet_wrap(~candidate_name)

```
Based on @fig-pollsamplesize-canidate-pct, there does not seem to be a relationship between the sample size of a poll the percentage of a candidate.

```{r}
#| label: fig-pollscore-canidate-pct
#| fig-cap: More Reliable Pollsters Score Higher For Trump
#| echo: false
#| warning: false

cleaned_data %>% 
  filter(!is.na(pollscore) & !is.na(pct)) %>% 
  ggplot(aes(x = pollscore, y = pct)) +
  geom_point(alpha = 0.4) + 
  geom_smooth(method = "lm", se = FALSE, color = light_red) +
  labs(x = "Pollscore",
       y = "Candidate Percentage") + facet_wrap(~candidate_name)


```
Figure @fig-pollscore-candidate-pct depicts the relationship between a pollster's poll score and a candidate's percentage. For Donald Trump, a negative correlation is observed, indicating that pollsters with lower poll scores tend to assign him a higher percentage of support. Conversely, Kamala Harris shows the opposite trend: as the poll score decreases, her percentage tends to rise. This suggests that more reliable polls, characterized by lower poll scores, report higher support for Trump compared to less reliable pollsters.

Based on @tbl-state-candidate-support, state might have to be a removed variable.

```{r}
#| label: fig-polldate-canidate-pct
#| fig-cap: Both Canidate Percentages Are Gaining Support Closer to the Election
#| echo: false
#| warning: false

cleaned_data %>% 
  filter(!is.na(days_taken_from_election) & !is.na(pct)) %>% 
  ggplot(aes(x = days_taken_from_election, y = pct)) +
  geom_point(alpha = 0.4) + 
  geom_smooth(method = "lm", se = FALSE, color = light_red) +
  labs(x = "Timing of Polls Relative to Election Day (Days Prior to the Election)",
       y = "Candidate Percentage") + facet_wrap(~candidate_name)

```
Based on @fig-polldate-canidate-pct, Trump and Harris are getting more votes in polls that are done closer to the election. This is a result of non-major candidate support rapidly decreasing. Specifically, Trump support is increasing at a faster rate than Harris. However, the polls in general show a slight lead for Harris throughout the last 100 days. 

The rapid decline in third party support could be due to Robert F. Kennedy dropping out of the race (this sentences should probably be in results or discussion section).

\newpage
# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.  

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
us


### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.


# Results

Our results are summarized in @tbl-modelresults.




# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}


# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows... 

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

# FiveThirtyEight Licenses
[FiveThirtyEight's data sets](https://github.com/fivethirtyeight/data/tree/master/polls) are used and modified by us under the [Creative Commons Attribution 4.0 International License](https://creativecommons.org/licenses/by/4.0/).

# Overview of American Electon System

# Brief Description of American Federal Goverment

The American federal or national government is split into three branches, executive., legislative, and judical The executive branch includes the president and the military. The legislative branch include two subgroups, the House of Representatives and the Senate. These two subgroups create laws. Every state has two Senators and one Representative per approximately 750,000 people. The judicial branch is court system.

## What is the Electoral College {#sec-electoral-college}

The Electoral College is the system used in the United States to elect the president and vice president. Instead of a direct popular vote, each state is allocated a certain number of electors based on its representation in Congress (the total of its Senators and Representatives). When voters cast their ballots, they are actually voting for a slate of electors pledged to a candidate. The candidate who receives a majority of electoral votes (270 out of 538) wins the presidency. This system means that winning the popular vote in a state generally results in winning all of that state's electoral votes. The only exception are the states of Maine and Nebraska, who award electoral votes by congressional district, with two additional votes given to the statewide winner.

The Electoral College results in some states being unnecessary to campaign in, as their strong historical voting pasterns towards either Democrats or Republicans make them unlikely to change, regardless of campaign efforts. Therefore, for statisticians, polling information from these states may not be that useful when trying to predict the outcome of an election. On the other hand, states that can vote either Democratic or Republican (swing states) are immensely important when predicting an election. As a result, campaigns spend hundreds of millions of dollars campaigning and understanding voters there.

```{r}
#| label: fig-currentstateofelection
#| fig-cap: 2024 U.S. Presidential Election State Forecast Map
#| echo: false
#| warning: false

# Define your colors for each state (replace these colors with your desired ones)
state_colors_legend <- c(
  "Republican" = dark_red,      # Dark Red
  "Lean Republican" = light_red, # Light Red
  "Toss-Up" = purple,         # Purple
  "Lean Democrat" = light_blue,   # Light Blue
  "Democrat" = dark_blue,        # Dark Blue
  "Split By District" = "gray"
)

state_colors <- c(
  alabama = dark_red,
  arizona = purple,
  arkansas = dark_red,
  california = dark_blue,
  colorado = dark_blue,
  connecticut = dark_blue,
  delaware = dark_blue,
  `district of columbia` = dark_blue,
  florida = light_blue,
  georgia = purple,
  idaho = dark_red,
  illinois = dark_blue,
  indiana = dark_red,
  iowa = light_red,
  kansas = dark_red,
  kentucky = dark_red,
  louisiana = dark_red,
  maine = "gray",
  maryland = dark_blue,
  massachusetts = dark_blue,
  michigan = purple,
  minnesota = light_blue,
  mississippi = dark_red,
  missouri = dark_red,
  montana = dark_red,
  nebraska = "gray",
  nevada = purple,
  `new hampshire` = light_blue,
  `new jersey` = dark_blue,
  `new mexico` = light_blue,
  `new york` = dark_blue,
  `north carolina` = purple,
  `north dakota` = dark_red,
  ohio = light_red,
  oklahoma = dark_red,
  oregon = dark_blue,
  pennsylvania = purple,
  `rhode island` = dark_blue,
  `south carolina` = dark_red,
  `south dakota` = dark_red,
  tennessee = dark_red,
  texas = light_red,
  utah = dark_red,
  vermont = dark_blue,
  virginia = light_blue,
  washington = dark_blue,
  `west virginia` = dark_red,
  wisconsin = purple,
  wyoming = dark_red
)

# Load the states map data
states_map <- map_data(map = "state")

state_colors <- setNames(state_colors, tolower(names(state_colors)))

color_to_status <- setNames(names(state_colors_legend), state_colors_legend)

states_map <- states_map %>%
  mutate(status = color_to_status[state_colors[region]])

plotclr <- c(dark_red, light_blue, purple, gray, light_blue, dark_blue)
legend_description <- c(
  "Likely Republican",
  "Lean Republican",
  "Swing States",
  "Lean Democrat",
  "Likely Democrat",
  "Split by District"
)


ggplot(states_map, aes(x = long, y = lat, group = group, fill = status)) +
  geom_polygon(color = "white", size = 0.2) +
  coord_map("albers", lat0 = 39, lat1 = 45) +
  scale_fill_manual(
    values = state_colors_legend,
    name = "Election Status"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10),
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank()
  )

```
The state statuses presented in this map are based on evaluations from @cnn, @foxnews, and @msnbc, which are generally agreed upon within the American political community. These organizations assessed historical voting patterns and recent polling data to derive their conclusions.

Notably, Nebraska and Maine are indicated in gray due to their delegates being split by district. In Nebraska, the state overall is projected to lean Republican, with the first and third districts also strongly favoring Republican candidates, while the second district leans Democratic. In Maine, the overall expectation is a Democratic leaning, consistent with its first district, though the second district leans Republican. Furthermore, Alaska and Hawaii are not in the map. Alaska is strongly favoring Republicans while Hawaii strongly favors democrats.

The seven generally agreed upon swing states are Pennsylvania, North Carolina, Georgia, Arizona, Nevada, Wisconsin, and Michigan with Texas, Florida, Nebraska District Two, Maine District 2, and Minnesota as the next closest races.

# States Poll Count

```{r}
#| label: tbl-state-polls-count
#| tbl-cap: Polls included in Analysis Per State
#| echo: false
#| warning: false

states <- c(state.name, "National")
polls_per_state_tbl <- tibble(state = states)

polls_per_state <- cleaned_data %>% group_by(state) %>% 
  summarise(polls_count = n_distinct(poll_id))

polls_per_state_tbl <- polls_per_state_tbl %>% 
  left_join(polls_per_state, by = "state") %>% 
  mutate(polls_count = if_else(!is.na(polls_count), polls_count, 0)) %>% 
  select(state, polls_count)

polls_per_state_tbl %>% kable(col.names = c("State", "Number of Polls In Analysis"))


num_states_missing <- count(polls_per_state_tbl %>% 
                              filter(polls_count == 0) %>% 
                              select(state))

```

\newpage


# References


